---
title: "Youtube Trending Videos"
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: Background

Often times, when I'm browsing on Youtube, I'm always wondering why those videos can jump to the trending list. There is a pretty handy and complex algorithm behind the scene.

However, for this analysis, I will focus on using statistical methods to solve three problems:

* Should YouTubers publish their videos on weekend or weekdays?

* Whether getting into trending required same amount of views among different groups? or are they different?

* Is there a relationship between views and number of tags when a video goes on trending?

### 1. Load the data and import packages

```{r, message=FALSE}
library(tidyverse)
library(lubridate)
youtube_orig <- read.csv("/Volumes/GoogleDrive/Shared drives/Statistics Final Presentation/youtube/youtube1.csv")
youtube <- youtube_orig
```

### 2. Data Overview

```{r}
rmarkdown::paged_table(head(youtube))
```

This youtube dataset includes data of trending videos from 12/01/2017 to 05/31/2018.

There are **16** variables in total:

* Numerical Variables: : views, likes, dislikes, comment_count

* Categorical Variables: channel_title, category_id, tags, comments_disabled, ratings_disabled, video_error_or_removed

* Time Variables: trending_date, publish_time

* Others: video_id, title, thumbnail_link, description

Each video may be put on the trending list more than once.


### 3. Data Processing

Since each video can show up several times in this dataset, lots of repetitive information exists. 

Moreover, all the numeric variables are cumulative, which means certain data processing methods are needed for the growth analysis concerning these variables.

I will select the first trending data for each video in order to avoid time-series influence

```{r}
youtube_first_trending <- youtube %>%
  group_by(video_id) %>%
  mutate(first_trending = min(ydm(trending_date))) %>%
  filter(ydm(trending_date) == first_trending) %>%
  separate(publish_time, c("date","time"), sep="T") %>%
  filter(date>="2017-12-01") %>%
  dplyr::select(video_id, trending_date, date, time,title, channel_title, 
                category_id, views,likes, dislikes, comment_count,tags) %>%
  distinct()
```

## Part 2: Two-sample T-Test

### 1. Video Views on Weekends vs. Weekdays

Now, we will start our statistical analysis on the youtube trending video datasets.

As a YouTuber, you will probably wonder if you should publish your videos during the weekends and weekdays to get more views. I have the intuition that people will tend to watch YouTube videos on weekends more than on weekdays.

Here comes our hypothesis:

* H0: Trending videos have the same number of views on weekends and weekdays.

* H1: Trending videos have different numbers of views ib weekebds and weekdays.

```{r}
youtube_ttest <- youtube_first_trending %>%
  mutate(week_day = weekdays(ymd(date))) %>%
  mutate(weekend=ifelse((week_day=="Saturday" | week_day=="Sunday"),1,0))

t.test(youtube_ttest$views ~ youtube_ttest$weekend,
       alternative = "two.sided", var.equal = TRUE)
```

### 2. Findings

Based on the t-test output, we can see we have a p-value of 0.566, which is far larger than 0.05, the pre-decided significance level. 

Therefore, the H0 cannot be rejected, and trending videos tend to have the same number of views during weekday and weekend.

This gives us the information that there is no difference in terms of views whether you decide to publish videos on weekdays and weekends.

## Part 3: ANOVA Test

### 1. Video Views and Category ID

```{r}
length(unique(youtube$category_id))
```

There are 16 video categories in our dataset. I'm curious about if there is a significant difference in views among different video categoris. Whether getting into Trending required same amount of views among different groups? or are they different?

* H0: Trending videos have the same number of views across different categories.

* H1: Trending videos have different numberd of views across different categories.

I did an ANOVA test here to compare the means of views when videos go on trending first time for each category.

```{r}
categoryTest <- aov(data = youtube_first_trending, views~as.factor(category_id))
summary(categoryTest)
TukeyHSD(categoryTest)
```

### 2. Findings

The result of the ANOVA test shows a p-value that is greatly smaller than 0.05, the pre-decided significance level. 

Therefore, the H0 should be rejected, and trending videos tend to have different numbers of views across different categories.

Moreover, the TukeyHSD test indicates the exact differences in video views between pairwise groups. `Category_10 (music)` tends to have the most number of views, while `category_25 (news & politics)` tends to have the least.

## Part 4: Regression Model

### 1. Relationship between video views and number of tags

YouTube videos usually have a lot of tags. The tags are separate by the "|". Let's see two examples below:

```{r}
head(youtube$tags,2)
```

Is there a relationship between views and number of tags when a video goes on trending?

In theory, the more tags a video has, the more likely it will show up in searches, and more people will watch it.

Our hypothesis: *number of tags has a positive relationship with views for videos first on trending*

Let's run a simple linear regression to test our findings:

\[Views\ = \beta_0\ + \beta_1\ * Num\ of\ Tags\]
\[H_0: \beta_1\ = 0\] 
\[H_1: \beta_1\ != 0\]

```{r}
youtube_tags <- youtube_first_trending %>% 
  mutate(tags = as.character(tags)) %>% 
  mutate(numTags = ifelse((tags == "[none]"), 
                          0, str_count(tags, "\\|")+1))

model_tags <- lm(data = youtube_tags, views~numTags)
summary(model_tags)
par(mfrow = c(2, 2))
plot(model_tags)
```

### 2. Findings

The coefficient has a 0.011 p-value that is smaller than 0.05, the pre-decided significance level. Thus, the H0 should be rejected, and the number of tags for a video can positively influence its number of views.

However, the following diagnostic plots show that the residuals are not normally distributed, and the r-square is only 0.001261. The F-statistic is 6.469, which also means this model is not a good fit to the dataset.

## Part 5: Revised Regression Model

### 1. Use music and news&politics categories as dummy variables

Can we improve our model? According to the ANOVA test in part 3, we know that category 10 (music) and category 25 (news & politics)'s views usually different from others. So I want to use *Music* and *News&Politics* as dummy variables and add to our linear model.

\[Views\ = \beta_0\ + \beta_1\ * Num\ of\ Tags\ + \beta_2\ * Dummy_{music}\ +\beta_3\ *\ Dummy_{news}\]
\[H_0: \beta_1\ = 0,\ \beta_2\ = 0,\ \beta_3\ = 0\]
\[H_1: \beta_1\ != 0,\ \beta_2\ != 0,\ \beta_3\ != 0\]

```{r}
youtube_dummy <- youtube_tags %>% 
  mutate(music = ifelse(category_id == 10, TRUE, FALSE)) %>% 
  mutate(news = ifelse(category_id == 25, TRUE, FALSE))

model <- lm(data = youtube_dummy, views~numTags+music+news)
summary(model)
par(mfrow = c(2, 2))
plot(model)
```

### 2. Findings

All three coefficients have p-values that are smaller than 0.05, the pre-decided significance level. Thus, the H0 should be rejected, and the number of tags and the music dummy has a positive impact on video views, while the news dummy has a negative one.

However, the following plots show that residuals are not normally distributed, and dummy variables cause a blank space in fitted values. The r-square is increased to 0.2879, and the F-statistic becomes 50.59, which means this revised model is a better fit than the original one.

## Part 6: Limitations

1. This dataset doesn't include data of those videos falling out of the trending list, which can be utilized to determine the views threshold of the trending list.

2. This analysis assumes the trending list is daily updated, while it actually may be updated every 15 minutes. This proxy leads to inaccuracy.

3. The revised model still doesn't have a high r-square and F-statistic, which means further analysis is needed to explore the relationship.

